Demes: a specification for interchanging complex demography

### 2021-06-29 paper.tex
Trim out spec-like details from popgen model.

In Demes, demographic models consist of one or more interacting populations (or
``demes"; to avoid confusion with the name of the specification itself we will
use the term ``population'' in this discussion, in the understanding that the
terms are interchangeable for our purposes). Populations are defined as a
collection of exchangable individuals that exist for a specified period of
time, following a well-defined set of rules regarding population sizes, mating
systems, etc. Ancestor/descendant relationships between populations can be
defined, as well as continuous or instantaneous migration between coexisting
populations.

Population and event times are written as units in the past, so that time zero
corresponds to the final generation or ``now'', and event times in the past are
values greater than zero with larger values for events that occur in the more
distant past. While a natural specification for time units is in generations,
other time units are allowed, such as years. In the case that the time units
are not given in generations, the generation time must also be specified so
that times can be converted into generations. In general, as time flows from
the past to the present, populations, epochs, and migration events should be
specified in their order of appearance, so that their times are in descending
order.

Models include one or more populations that have sizes greater than zero for
the duration of their existences, and population sizes are given as the number
of individuals. A population must exist for some duration of time greater than
zero, so that its start time must be larger than its end time. It may have one
or more ancestors, which are other populations that exist at the population's
start time. If one ancestor is specified, the first generation is constructed
by randomly sampling parents from the ancestral population to contribute to
offspring in the newly generated population. If more than one ancestor is
specified, the proportions of ancestry from each contributing population must
be provided, and those proportions must sum to one. In this case, parents are
chosen randomly from each ancestral population with probability given by those
proportions. If no ancestors are specified, the population is assumed to have
start time equal to infinity, as individuals are not assumed to spontaneously
spring into existence.

Within populations, population sizes and mating systems are specified within
epochs. Epochs are non-overlapping intervals of time that cover the entire
existence interval of the population. Each epoch specifies the population size
over that interval, which can be a constant value or function defined by start
and end sizes that must remain positive.  If an epoch has a start time of
infinity, the population size for that epoch must be constant. Epochs can also
specify parameters for nonrandom mating, such as selfing or cloning rates,
which give the probability that offspring are generated from one generation to
the next by self-fertilization or cloning of an individual \textbf{need to
precisely specify once this is determined}.

Finally, individuals in a population may have parents from a different
population through migrations. These can be defined as continuous migration
rates over time intervals for which populations coexist or through
instantaneous (or pulse) migration events at a given time. Continuous migration
rates are defined as the probability that parents in the ``destination''
population are chosen from the ``source'' population.  Migration rates are thus
per generation and must be less than one. Furthermore, if more than one source
population have continuous migration into the same destination population, the
sum of those migration rates must also be less than one, as rates define
probabilities. The probability that parents come from the same population is
just one minus the sum of incoming migration rates. On the other hand, pulse
migration events specify the instantaeous replacement of a given fraction of
individuals in a destination population by individuals with parents from a
source population.  The fraction must be between zero and one, and if more than
one pulse occurs at the same time, those replacement events are applied
sequentially in the order that they are specified in the model.

>>>>

In Demes, demographic models consist of one or more interacting populations (or
``demes"; to avoid confusion with the name of the specification itself we will
use the term ``population'' in this discussion, in the understanding that the
terms are interchangeable for our purposes). Populations are defined as a
collection of exchangeable individuals that exist for a specified period of
time, following a well-defined set of rules regarding population sizes, mating
systems, etc. Ancestor/descendant relationships between populations can be
defined, as well as continuous or instantaneous migration between coexisting
populations.

Population and event times are written as units in the past, so that time zero
corresponds to the final generation or ``now'', and event times in the past are
values greater than zero with larger values for events that occur in the more
distant past. While a natural specification for time units is in generations,
other time units are allowed, such as years.
\jkcomment{Why did we choose time to flow in this direction?}

Population sizes are given as the number of individuals.
\jkcomment{As opposed to genomes. Why did we do this?}
Sizes and mating system details are specified for each population within
epochs. Epochs are contiguous time intervals that define
the existence interval of the population. Each epoch specifies the population size
over that interval, which can be a constant value or function defined by start
and end sizes that must remain positive.

Within a population, individuals are assumed to be exchangeable, but
there are also parameters for nonrandom mating, such as selfing or cloning rates,
which give the probability that offspring are generated from one generation to
the next by self-fertilization or cloning of an individual
\textbf{need to precisely specify once this is determined}.
\jkcomment{More detail here. Also, why only selfing and cloning, why not selection
coefficients and whatnot?}

A population may have one
or more ancestors, which are other populations that exist at the population's
start time. If one ancestor is specified, the first generation is constructed
by randomly sampling parents from the ancestral population to contribute to
offspring in the newly generated population. If more than one ancestor is
specified, the proportions of ancestry from each contributing population must
be provided, and those proportions must sum to one. In this case, parents are
chosen randomly from each ancestral population with probability given by those
proportions.

Finally, individuals in a population may have parents from a different
population through migrations. These can be defined as continuous migration
rates over time intervals for which populations coexist or through
instantaneous (or pulse) migration events at a given time. Continuous migration
rates are defined as the probability that parents in the ``destination''
population are chosen from the ``source'' population.
On the other hand, pulse
migration events specify the instantaneous replacement of a given fraction of
individuals in a destination population by individuals with parents from a
source population.
